---
title: "Differential Host Competence of Amphibian Communities for Ranavirus"
output:
  html_document:
    df_print: paged
---

This script is an attempt to bring together the relevant information and scripts from my work up until now. An effort to organize, scrutinize, and finalize. This is formatted to act as a proposal for bringing the project to completion


```{r,load packages, include = FALSE}
library(tidyverse)
library(magrittr)
library(deSolve)
library(knitr)
library(ggridges)
library(reshape2)
```

This project can be split up into three distinct sections
 (A) Data exploration of Ranavirus prevalence in wetlands of amphibians
 (B) Metacommunity analysis of wetlands of amphibians and their propensity for parasite invasion/prevalence
 (C) Theoretical model of transmission for ranavirus in a mixed-host variable-competency community




Composition of hosts within a community plays a role in transmission potential of an introduced parasite. The competency of hosts is an important mechanism behind the relationship between community composition and parasite success. Using empirical data and a theoretical model for disease transmission, we can hypothesize the effects of community composition in a real system. The system being considered is Ranavirus in amphibian communities. Ranavirus, a dsDNA virus that mostly affects amphibians, reptiles, and fish is a known contributor to amphibian declines throughout the globe. Hosts vary in quality for transmission, and environmental transmission is known to occur in addition to contact transmission. These characteristics make this system an excellent opportunity to study the effects of community composition on disease transmission and should allow us to ask further questions about the biological realism that is relevant for studying this relationship. 


Ranavirus Summary
```{r, Ranavirus Summary}
data <- read_csv("competence.csv")

#Plot of RV status over time
rv <- data %>% select(Month,WetAltID,RV.Status)
rv$Month <- factor(data$Month,levels=c("Feb","Mar","Apr","May","Jun","Jul"))
rv %>% 
  filter(RV.Status==1) %>% 
  ggplot(., aes(x=Month,
                y=WetAltID,
                group=WetAltID,
                fill=as.factor(WetAltID))) + 
  geom_density_ridges_gradient(show.legend=F) +
  labs(title="# of RV cases per site over time")




#looking at rv of species and drawing a line to divide "low" and "high" competent
rv_species <- data %>% select(vl.4:vl.42) %>% distinct()
rv_species %<>% pivot_longer(.,vl.4:vl.42) %>% filter(value>0)
rv_species %<>% add_column(log_value = log(rv_species$value))
rv_species$name <- factor(rv_species$name, levels = rv_species$name[order(rv_species$log_value)])
rv_species$actual_name <- factor(c("N. viridescens", "S. holbrookii", "A. terrestris", "A. gryllus", "P. crucifer", "H. gratiosa", "H. avivoca",
                                   "P. nigrita", "P. ornata", "G. carolinensis", "L. clamitans", "L. sphenocephalus"))
rv_species$actual_name <- factor(rv_species$actual_name, levels = rv_species$actual_name[order(rv_species$log_value)])

#natural log of values
rv_species %>% ggplot(.,aes(x=actual_name,y=log_value))+geom_point() + xlab("Species") + ylab("ln(Viral Load)") + geom_hline(yintercept = log(20)) + labs(title = "Viral Loads") + theme_classic() + theme(axis.text.x = element_text(angle = 30, hjust = 1))


```
Same summary as before but with a different measure of competency. These measures are calculated using only RV positive individuals
```{r, Ranavirus Summary 2}
data <- read_csv("competence_rv_pos.csv")

#Plot of RV status over time
rv <- data %>% select(Month,WetAltID,RV.Status)
rv$Month <- factor(data$Month,levels=c("Feb","Mar","Apr","May","Jun","Jul"))
rv %>% 
  filter(RV.Status==1) %>% 
  ggplot(., aes(x=Month,
                y=WetAltID,
                group=WetAltID,
                fill=as.factor(WetAltID))) + 
  geom_density_ridges_gradient(show.legend=F) +
  labs(title="# of RV cases per site over time")




#looking at rv of species and drawing a line to divide "low" and "high" competent
rv_species <- data %>% select(vl.9:vl.42) %>% distinct()
rv_species %<>% pivot_longer(.,vl.9:vl.42) %>% filter(value>0)
rv_species %<>% add_column(log_value = log(rv_species$value))
rv_species$name <- factor(rv_species$name, levels = rv_species$name[order(rv_species$log_value)])
rv_species$actual_name <- factor(c("N. viridescens", "S. holbrookii", "A. terrestris", "P. crucifer", "H. gratiosa", "H. avivoca",
                                   "P. nigrita", "P. ornata", "G. carolinensis", "L. clamitans", "L. sphenocephalus"))
rv_species$actual_name <- factor(rv_species$actual_name, levels = rv_species$actual_name[order(rv_species$log_value)])

#natural log of values
rv_species %>% ggplot(.,aes(x=actual_name,y=log_value))+geom_point() + xlab("Species") + ylab("ln(Viral Load)") + geom_hline(yintercept = 5) + labs(title = "Viral Loads") + theme_classic() + theme(axis.text.x = element_text(angle = 30, hjust = 1))


```


Select only site-level data for absolute abundance (species included are only those for which viral load was included)
```{r, Absolute Abundance}
site_time_cc_tot <- data %>% select(WetAltID, Month.1, AB2:AB42, cc) %>% select(-ABAmb,-ABSal) %>% distinct()

site_time_cc_tot <- mutate(site_time_cc_tot, ABHigh = (AB26 + AB42 + AB21 + AB9))
site_time_cc_tot <- mutate(site_time_cc_tot, ABLow = (AB38 + AB34 + AB41 + AB35 + AB20 + AB28 + AB24))
site_time_cc_tot <- mutate(site_time_cc_tot, total = ABHigh + ABLow)

#how to facet this to look at all sites?
site_time_cc_tot %>% filter(WetAltID==8) %>% select(Month.1, ABHigh, ABLow) %>% melt() %>%
  ggplot(., aes(y = value, x = Month.1, fill = variable)) +
  geom_bar(position = "stack", stat = "identity")  +
  labs(title = "Absolute Abundace for high and low competence at wetland 8")
```

Absolute abundance (including all species i.e. more species included in ABLow)
```{r, Absolute Abundance 2}
site_time_cc_tot <- data %>% select(WetAltID, Month.1, AB2:AB42, cc) %>% select(-ABAmb,-ABSal) %>% distinct()

site_time_cc_tot <- mutate(site_time_cc_tot, ABHigh = (AB26 + AB42 + AB21 + AB9))
site_time_cc_tot <- mutate(site_time_cc_tot, 
                           ABLow = (AB2 + AB3 + AB4 + AB5 + AB6 + AB8 + AB20 + AB24 + AB27 + AB28 + AB29 + AB31 + AB34 + AB35 + AB38 + AB39 + AB41))
site_time_cc_tot <- mutate(site_time_cc_tot, total = ABHigh + ABLow)

#how to facet this to look at all sites?
site_time_cc_tot %>% filter(WetAltID==8) %>% select(Month.1, ABHigh, ABLow) %>% melt() %>%
  ggplot(., aes(y = value, x = Month.1, fill = variable)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(title = "Absolute Abundace for high and low competence at wetland 8 (includes all species)")
```

Select only site-level data for relative abundance
```{r, Relative Abundance}
site_time_cc <- data %>% select(WetAltID, Month.1, RA2:RA42, cc) %>% select(-RAAmb,-RASal) %>% distinct()

site_time_cc <- mutate(site_time_cc, RAHigh = (RA9 + RA21 + RA26 + RA42))
site_time_cc <- mutate(site_time_cc, 
                       RALow = (RA2 + RA3 + RA4 + RA5 + RA6 + RA8 + RA20 + RA24 + RA27 + RA28 + RA29 + RA31 + RA34 + RA35 + RA38 + RA39 + RA41))
site_time_cc <- mutate(site_time_cc, total = RAHigh + RALow)

#how to facet this to look at all sites?
site_time_cc %>% filter(WetAltID==8) %>% select(Month.1, RAHigh, RALow) %>% melt() %>%
  ggplot(., aes(y = value, x = Month.1, fill = variable)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(title = "Relative Abundance of high and low competence for wetland 8")

site_time_cc %>% filter(WetAltID==8) %>%
  ggplot(., aes(y=cc, x=Month.1)) +
  geom_point() +
  labs(title = "CC over time for Wetland 8")

data %>% select(WetAltID, Month.1, Prevalence) %>% filter(WetAltID==8) %>%
  ggplot(., aes(y=Prevalence, x=Month.1)) +
  geom_point() +
  labs(title = "Prevalence over time for Wetland 8")
```

I start with a summary of the ranavirus dataset that looks at the variance in ranavirus prevalence throughout a community over space and time and the viral loads of different amphibian species. The viral loads of these species provides us with insight about how to conceptualize these communities in a simplifying way. A simpler way to think about these systems is to split them up into two species with different levels of competency - one high and one low. This framework allows for a manageable transmission model with the ability to alter characteristics such as community composition, viral characteristics, and species characteristics.


Comparing prevalence and cc for sites and time
```{r, Prevalence and Community Competence}
data <- read_csv("competence.csv")

#dataset for looking at prevalence and cc for sitextime
prev_cc <- data %>% select(WetAltID, Month.1, cc, Prevalence, Month)
prev_cc$Month.1 <- factor(data$Month,levels=c("Feb","Mar","Apr","May","Jun","Jul"))
prev_cc %<>% distinct() %>% distinct() %>% remove_missing()

#heatmap of cc for sitextime
ggplot(prev_cc, aes(x=Month.1, y=WetAltID)) + geom_tile(aes(fill=cc)) + labs(title = "Community Competence", x="Month", y = "Wetland") + scale_fill_gradient(low = "white",high = "indianred1")

#heatmap of prevalence for sitextime
ggplot(prev_cc, aes(x=Month.1, y=WetAltID)) + geom_tile(aes(fill=Prevalence)) + labs(title = "Prevalence", x = "Month", y = "Wetland") + scale_fill_gradient(low = "white",high = "indianred1")


#plot cc by Prevalence(this is %prevalence)
prev_cc %>% ggplot(.,aes(x=cc, y=Prevalence)) + 
  geom_point() +
  theme_classic() +
  labs(title = "without lag")

#put everything in order by wetland and month
order <- prev_cc[order(prev_cc$WetAltID, prev_cc$Month.1),]

#this removes months out of sequence
order <- order[-c(4,37,38,45,85),]

#create new column that includes previous month's value for cc
for(n in 2:86){
order$lag[n] <- order$cc[n-1]
}

#plot cc by prevalence with lag
order %>% ggplot(.,aes(x=lag, y=Prevalence)) + 
  geom_point() +
  theme_classic() +
  labs(title = "with lag")

#remove the first entry for each wetland to remove the carryover from the last wetland
clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)

#plot cleaned plot with lag
clean %>% filter(Prevalence>0) %>% ggplot(.,aes(x=lag, y=Prevalence)) +
  geom_point() +
  theme_classic() + geom_smooth(method = "lm") +
  labs(title = "", x = "Community Competence", y = "Prevalence")
clean2 <- clean %>% drop_na()

#correlation test
cor.test(clean2$lag,clean2$Prevalence,method="spearman")
```

Community competence is a measure for the overall competency of a community as opposed to individual values of competency for each species of host. This metric is calculated using the relative abundance of a species multiplied by its competency value. This provides a sum for all species within a community and the competency that they contribute to that system. Community competence is a derived characteristic that agrees with our assumption of host competency playing a role in the transmission potential of an introduced parasite. An empirical measure of transmission potential might be the prevalence of ranavirus at a site. We can compare our derived measure and this empirical measure by plotting their correlation (i.e. higher competence relates to higher prevalence), and we consider a time lag of 1 month between an observation of competence and an observation of prevalence. When we do this, a positive relationship appears confirming our hypothesis that community competence correlates with prevalence.


Basic function for developing contour plots using simulated data
```{r, Invasion}
contour <- function(trans_1_min, trans_1_max, trans_3_min, trans_3_max, mort1, mort2, degr) {
data <- expand_grid("trans1" = seq(from = trans_1_min, to = trans_1_max, by = 0.0001),
                    "trans2" = c(0.0001),
                    "trans3" = seq(from = trans_3_min, to = trans_3_max, by = 0.0001),
                    "shed1" = c(rep(1/2, length("trans1"))),
                    "shed2" = c(rep(1/2, length("trans1"))),
                    "reco1" = c(rep(1/10, length("trans1"))),
                    "reco2" = c(rep(1/10, length("trans1"))),
                    "mort1" = c(mort1), 
                    "mort2" = c(mort2), 
                    "degr" = c(degr), 
                    "birth1" = c(rep(5/3, length("trans1"))),
                    "birth2" = c(rep(5/3, length("trans1"))))

data %<>% add_column(dfe1 = (data$birth1/data$mort1),dfe2 = (data$birth2/data$mort2)) 

mat <- matrix(1, nrow = 3, ncol = 3)


m <- nrow(data)

for(n in 1:m){
  mat[1] = (data$trans1[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
  mat[2] = (data$trans1[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
  mat[3] = (data$shed1[n]/data$degr[n])
  mat[4] = (data$trans2[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
  mat[5] = (data$trans2[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
  mat[6] = (data$shed2[n]/data$degr[n])
  mat[7] = (data$trans3[n]*data$dfe1[n])/(data$reco1[n]+data$mort1[n])
  mat[8] = (data$trans3[n]*data$dfe2[n])/(data$reco2[n]+data$mort2[n])
  mat[9] = 0
  eigen <- max(eigen(mat)$values)
  data$eigen[n] <- eigen
}

data$span1 <- 1/data$mort1
data$span2 <- 1/data$mort2

data$epid <- data$eigen > 1
data$prop1 <- data$trans1/data$trans2
data$prop3 <- data$trans3/data$trans2
data$prop_mort <- data$mort1/data$mort2
data$cap1 <- data$birth1/data$mort1
data$cap2 <- data$birth2/data$mort2
data$cap_prop <- data$cap1/data$cap2
data$tot <- data$cap1 + data$cap2
data$cc <- data$trans1*(data$cap1/data$tot) + data$trans2*(data$cap2/data$tot)
plots <- data %>% select(trans1, trans2, prop1, prop3, trans3, eigen, prop_mort, mort1, mort2, degr, cap1, cap2, cap_prop, tot, cc) %>% distinct()
return(plots)
}

#"default" parameters
plotr <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/50,mort2 = 1/50,degr = 1/2)

plotr %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort2), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(values = c(NA, "red")) + 
  labs(x = "Environmental", y = "Contact") +
  xlim(2,10) +
  ylim(2,10)
```

Use this to more easily make all the plots for varying different parameters
```{r, Changing Parameters}
plot_mort1 <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/50,1/60,1/70),mort2 = 1/50,degr = 1/2)

plot_mort1 %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(values = c(NA, "red")) + 
  labs(x = "Environmental", y = "Contact") +
  xlim(2,10) +
  ylim(2,10) + 
  labs(title = "vary mort1")

plot_mort2 <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/50,mort2 = c(1/50,1/60,1/70),degr = 1/2)

plot_mort2 %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort2), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(values = c(NA, "red")) + 
  labs(x = "Environmental", y = "Contact") +
  xlim(2,10) +
  ylim(2,10) + 
  labs(title = "vary mort2")

plot_degr <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/50,mort2 = 1/50,degr = c(1/2,1/3,1/4))

plot_degr %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=degr), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(values = c(NA, "red")) + 
  labs(x = "Environmental", y = "Contact") +
  xlim(2,10) +
  ylim(2,10) + 
  labs(title = "vary degr")

```
Assuming that host competency is a relevant metric for evaluating the likelihood of an invasion of a parasite, we can numerically derive R0 from a set of equations. In line with our hypothesis, a greater value of community competency results in a larger range of parameter values in which a parasite can invade. This is represented when comparing the inner lines for plots (vary mort1 and vary mort2). The inner lines correspond and have the same total size for the combined populations but the proportion of highly competent individuals is higher in the "vary mort1" plot. Additionally, we can vary the rate of degradation of the free-infectious virus and change the slope of this curve, altering the ability for environmental transmission to compensate for direct transmission. As we favor environmental transmission, the importance of community composition is reduced. Currently, both species are equally competent in their ability to become infected but they vary in their ability to infect susceptible individuals.


Transmission Dynamics

```{r, Transmission Dynamics}
rm0<-function(t,x,params){
  Sa <- x[1]  
  Ia <- x[2]
  Ra <- x[3]
  Sb <- x[4]  
  Ib <- x[5]
  Rb <- x[6]
  V <- x[7]
  with(as.list(params),{
    dSa <- -beta*Sa*Ia - betab*Sa*Ib - phi*Sa*V + lambda - mu*Sa
    dIa <- beta*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alpha*Ia - mu*Ia
    dRa <- alpha*Ia - mu*Ra
    
    dSb <- -betab*Sb*Ib - beta*Sb*Ia - phi*Sb*V + lambda - mu*Sb
    dIb <- betab*Sb*Ib + beta*Sb*Ia + phi*Sb*V - alpha*Ib - mu*Ib
    dRb <- alpha*Ib - mu*Rb
    
    dV <- sigma*Ia + sigma*Ib - epsilon*V
    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
    list(res)
  })}
maxTime <- 300.0 # time is in years - run model for this time
times<-seq(0,maxTime,by=1) # how long we run the model for
# notes on params
# mu =     <- density-independent mortality rate
# beta =       <- contact transmission rate by infected A
# betab =       <- contact transmission rate by infected B
# phi =      <- environmental transmission rate
# lambda =       <- density-dependent growth rate
# alpha =     <- recovery rate
# sigma =    <- shedding rate
# epsilon =       <- viral degradation rate
params<-c(mu=1/50,
          beta=0.001,
          betab=0.0001,
          phi=0.001,
          lambda=5/3,
          alpha=1/10,
          sigma=1/2,
          epsilon=1/2)  # model parameters

xstart<-c(Sa=116,
          Ia=1,
          Ra=0,
          Sb=83,
          Ib=1,
          Rb=0,
          V=0)  # initial conditions

output<-as.data.frame(lsoda(xstart,times,rm0,params)) # tells computer to solve (integrate) equations

output %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.5))+
  labs(y="N",x="Time",col="Population")+
  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland")+theme(plot.title = element_text(hjust=0, size=10))
```

Transmission Dynamics with greater heterogeneity in species characteristics

```{r, Transmission Dynamics 2}
rm0<-function(t,x,params){
  Sa <- x[1]  
  Ia <- x[2]
  Ra <- x[3]
  Sb <- x[4]  
  Ib <- x[5]
  Rb <- x[6]
  V <- x[7]
  with(as.list(params),{
    dSa <- -betaa*Sa*Ia - betab*Sa*Ib - phi*Sa*V + lambdaa - mua*Sa
    dIa <- betaa*Sa*Ia + betab*Sa*Ib + phi*Sa*V - alphaa*Ia - mua*Ia
    dRa <- alphaa*Ia - mua*Ra
    
    dSb <- -betab*Sb*Ib - betaa*Sb*Ia - phi*Sb*V + lambdab - mub*Sb
    dIb <- betab*Sb*Ib + betaa*Sb*Ia + phi*Sb*V - alphab*Ib - mub*Ib
    dRb <- alphab*Ib - mub*Rb
    
    dV <- sigma*Ia + sigma*Ib - epsilon*V
    res<-c(dSa,dIa,dRa,dSb,dIb,dRb,dV)
    list(res)
  })}
maxTime <- 300.0 # time is in years - run model for this time
times<-seq(0,maxTime,by=1) # how long we run the model for
# notes on params
# mu =     <- density-independent mortality rate
# beta =       <- contact transmission rate by infected A
# betab =       <- contact transmission rate by infected B
# phi =      <- environmental transmission rate
# lambda =       <- density-dependent growth rate
# alpha =     <- recovery rate
# sigma =    <- shedding rate
# epsilon =       <- viral degradation rate
params<-c(mua=1/40,
          mub=1/50,
          betaa=0.001,
          betab=0.0001,
          phi=0.001,
          lambdaa=10/3,
          lambdab=5/3,
          alphaa=1/15,
          alphab=1/10,
          sigma=1/2,
          epsilon=1/2)  # model parameters

xstart<-c(Sa=116,
          Ia=1,
          Ra=0,
          Sb=83,
          Ib=1,
          Rb=0,
          V=0)  # initial conditions

output<-as.data.frame(lsoda(xstart,times,rm0,params)) # tells computer to solve (integrate) equations

output %>% ggplot(.,aes(x=time))+
  geom_line(aes(y=Sa,col="Susceptible A"))+
  geom_line(aes(y=Ia,col="Infected A"))+
  geom_line(aes(y=Ra,col="Recovered A"))+
  geom_line(aes(y=Sb,col="Susceptible B"))+
  geom_line(aes(y=Ib,col="Infected B"))+
  geom_line(aes(y=Rb,col="Recovered B"))+
  geom_line(aes(y=V,col="Virus"))+
  scale_colour_manual(values = c("red","orange","yellow","green","blue","black","violet"))+
  theme(legend.position = c(0.85,0.8))+
  labs(y="N",x="Time",col="Population")+
  ggtitle("2-Species Transmission Model for Ranavirus in an Amphibian Wetland")+theme(plot.title = element_text(hjust=0, size=10))
```

These plots look at the transmission dynamics over time for a scenario where the demography of the species are equal and when they are different. In the scenario when they are different we observe a tradeoff between high competency and shorter lifespan and higher birth rate. But this does not consider pathogen-induced mortality which we might expect to be higher for the more competent species that invests less into immunity and dies quicker anyways. 



Further questions and directions for this project

 - How might space correlate to changes in community competence or disease prevalence?
     - Can we map out the sites and their distances to each other to see if proximity to highly infected/competent sites are increase the likelihood        of infection/community competence in another?
 - How to apply this model to our data to observe the role that community competence plays in each site?
 - How to consider pathogen-induced mortality and the tradeoffs we might expect from an immunity-reproduction relationship?
 - How to observe community assembly and disassembly over time and the changing risks for disease?



This chunk creates a spatial matrix for all sites and their distances from each other
```{r}
##Measuring distance between points
library(geosphere)
library(biogeo)
library(raster)

latlon <- read_csv("srs_wetland_coordinates.csv")

latlon$dd_x <- dms2dd(latlon$xdd,latlon$xmm,latlon$xss,latlon$xdir)
latlon$dd_y <- dms2dd(latlon$ydd,latlon$ymm,latlon$yss,latlon$ydir)

srs_lat_lon <- tibble(WetAltID = latlon$Wetland, x = latlon$dd_x, y = latlon$dd_y)
#srs_lat_lon$latlon <- paste(srs_lat_lon$x,",", srs_lat_lon$y)

distances <- pointDistance(c(srs_lat_lon$x[1],srs_lat_lon$y[1]),c(srs_lat_lon$x[2],srs_lat_lon$y[2]), lonlat = TRUE)
matrix <- cbind(srs_lat_lon$x,srs_lat_lon$y)
dist_mat <- distm(matrix, fun = distGeo)
colnames(dist_mat) <- c(srs_lat_lon$WetAltID)
rownames(dist_mat) <- c(srs_lat_lon$WetAltID)


srs_lat_lon %>% ggplot(.,aes(x = x, y = y)) + geom_point() + geom_text(label = srs_lat_lon$WetAltID) 
dist_mat
```

Grab climate data from worldclim for each site
```{r}
r <- getData("worldclim",var="bio",res=10)

r <- r[[c(1,12)]]
names(r) <- c("Temp","Prec")

lats <- latlon$dd_x
lons <- latlon$dd_y

coords <- data.frame(x=lons,y=lats)

points <- SpatialPoints(coords, proj4string = r@crs)

values <- raster::extract(r,points)

climate <- cbind.data.frame(coordinates(points),values)
climate
```



This chunk creates a matrix for all amphibian pairwise phylogenetic distances and simulates a random walk for competence as a trait
```{r}
library(ape)
library(rotl)
library(phangorn)
library(geiger)
library(phytools)

#trait1 <- rTraitCont(tree, "OU", theta = 0, alpha=.1, sigma=.01)


names <- c("Notophthalamus viridescens", "Scaphiopus holbrookii", "Anaxyrus terrestris", "Pseudacris crucifer", "Hyla gratiosa", "Hyla avivoca",
                  "Pseudacris nigrita", "Pseudacris ornata", "Gastrophyrne carolinensis", "Lithobates clamitans", "Lithobates sphenocephalus",
           "Acris gryllus", "Hyla chrysoscelis", "Hyla cinerea", "Hyla femoralis", "Lithobates catesbeianus", "Siren intermedia")

resolved <- tnrs_match_names(names)

#tree <- tol_induced_subtree(resolved$ott_id)
#plot(tree, edge.width = "3")

#plot.phylo(tree, edge.width = '3', show.tip.label = TRUE)

species <- resolved$unique_name
species <- gsub(" ", "_", species)

amphibians <- read.nexus("asup_just_tree.txt")
#plot(amphibians, edge.width = "1")

str(amphibians)


setdiff(species,amphibians$tip.label)
species <- gsub("Dryophytes","Hyla",species)
species <- gsub("gratiosus","gratiosa",species)
species <- gsub("cinereus","cinerea",species)
species <- gsub("Anaxyrus","Bufo",species)
setdiff(species,amphibians$tip.label)

#prune
pruned.tree<-drop.tip(amphibians,amphibians$tip.label[-match(species, amphibians$tip.label)])

plot(pruned.tree)

#phylo distances
phy_dist <- cophenetic.phylo(pruned.tree)
phy_dist

#random walk for competence
trait1 <- rTraitCont(pruned.tree, "OU", theta = 0, alpha=.1, sigma=.01)
trait1 <- base::as.data.frame(trait1, row.names = NULL)
trait1 <- rownames_to_column(trait1, "species")
colnames(trait1) <- c("species", "value")
#plotted with squares to remove negative values
ggplot(trait1, aes(x = species, y = value)) + geom_point() + theme_classic() + theme(axis.text.x = element_text(angle = 30, hjust = 1))
```



